---
title: "Bubble plot"
format:
  html:
   fig-width: 10
   fig-height: 10
execute: 
  warning: false

---

A bubble plot is essentially a scatterplot where a third numeric variable determines the size of the circles.

A scatterplot illustrates the relationship between two numeric variables. Each dot represents an observation, with its position on the X (horizontal) and Y (vertical) axes indicating the values of the two variables.


## Table (Data format)

This is an example of a data format containing the necessary information for bubble plot visualizations. The table includes mock data specifically generated for this purpose.


```{r}

library(data.table)
library(reactable)

df <- fread("../inst/data/bubble_plot.txt") 


reactable(
  
  df, 
  defaultPageSize = 5,
  theme = reactableTheme(
    backgroundColor  = "transparent"
  )
)
```


## Bubble plot

The `ggplot2` package enables the creation of bubble charts using the `geom_point()` function.

This plot contains different pathways. The size of the bubble depends on the count and the color of the bubble depends on the -log(p.adj).

```{r}

# libraries -------
library(ggplot2)

ggplot(data = df) +
  
  geom_point(
    aes(x = NES, y = Description,
        size = `Count`, 
        fill = log_padj
        ),
    
    shape = 21, stroke = .25
  ) +
  
  scale_size_continuous(range = c(2, 7)) +
  
  scale_fill_gradient(
    low = "grey", high = "#f30000",
    guide = guide_colorbar(
      title = "-log10(p.adj)",
      barheight = unit(10, "lines"),
      barwidth = unit(.75, "lines")
    )
  ) +
  
  theme_minimal() +
  
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    
    panel.grid.major = element_line(linewidth = .4),
    
    panel.border = element_rect(fill = NA, linewidth = .4),
    axis.ticks = element_line(linewidth = .4),
    
    plot.margin = margin(20, 20, 20, 20)
  ) +
  
  facet_grid(
    rows = vars(Group),
    scales = "free_y",
    space = "free_y")

```
