---
title: "Diversity estimation"
---

Diversity estimation within immune cell populations is a fundamental analysis in immunoinformatics. Diversity estimation refers to the process of analyzing and quantifying the diversity of immune cells, such as T cells or B cells. 

## Diversity indeces
There are several approaches to the estimation of repertoire diversity implemented in the repDiversity function. The method parameter similarly to above mentioned functions sets the means for diversity estimation. You can choose one of the following methods:

- **shannon** Shannon index takes into account both the number of different species and the distribution of individuals among those species, providing a single numerical value that reflects the diversity within a community. Higher Shannon index values indicate greater diversity.
- **chao1** Chao is a nonparameteric asymptotic estimator of species richness (number of species in a population).
- **hill** Hill numbers are a mathematically unified family of diversity indices (differing only by an exponent q).
- **div** true diversity, or the effective number of types, refers to the number of equally abundant types needed for the average proportional abundance of the types to equal that observed in the dataset of interest where all types may not be equally abundant.
- **gini.simp** The Gini-Simpson index is the probability of interspecific encounter, i.e., probability that two entities represent different types.
- **inv.simp** Inverse Simpson index is the effective number of types that is obtained when the weighted arithmetic mean is used to quantify average proportional abundance of types in the dataset of interest.
- **gini** The Gini coefficient measures the inequality among values of a frequency distribution (for example levels of income). A Gini coefficient of zero expresses perfect equality, where all values are the same (for example, where everyone has the same income). A Gini coefficient of one (or 100 percents ) expresses maximal inequality among values (for example where only one person has all the income).
- **raref** Rarefaction is a technique to assess species richness from the results of sampling through extrapolation.

## Table (Data format)

This is an example of a data format containing the necessary information for diversity estimation visualizations. The table includes mock data specifically generated for this purpose.

The "Sample" column contains unique identifiers for each sample, while the "Group" column indicates the different groups to which the samples belong. The "Patient" column provides information about the respective patients associated with each sample. The "Shannon" column corresponds to the Shannon index utilized for the diversity analysis.

```{r}

library(data.table)
library(reactable)

df = fread("../inst/data/diversity1.txt")

reactable(
  
  df, 
  theme = reactableTheme(
    backgroundColor  = "#d9e3f1"
  )
)
```

## Boxplot

Boxplot is one of the most commonly used chart types for comparing the distribution of a numeric variable across multiple groups.

In a boxplot, the central line within the box represents the median of the data, dividing it into two equal parts. The edges of the box represent the upper and lower quartiles, while the extreme lines extend to the the highest and lowest values within the data range, excluding outliers.

Boxplots are created using the `geom_boxplot()` function from the `ggplot2` package.

### Boxplot 1

It's important to note that while boxplots provide a summary of data distribution for each group, they may hide the underlying distribution details. 

To address this concern, a common practice is to overlay individual data points using `geom_point()` behind the boxplot, offering a clearer visualization of the dataset's distribution.

```{r}
library(ggplot2)
library(ggforce)

library(ggnewscale)
library(paletteer)

library(colorspace)

# plot 1 ---------------------

df |>
    ggplot(aes(Group, Shannon)) +
    
    geom_point(
        aes(fill = Group),
        position = position_jitternormal(sd_y = 0, sd_x = .08),
        shape = 21, size = 2, stroke = .15, color = "white"
    ) +

    scale_fill_manual(values = paletteer_d("ggsci::hallmarks_light_cosmic")) +
    
    geom_boxplot(width = .15, outlier.shape = NA) +
    
    theme_minimal() +
    
    theme(
        legend.position = "none",
        
        axis.line = element_line(linewidth = .55),
        axis.ticks = element_line(linewidth = .55),
        
        panel.grid.major = element_line(linewidth = .55),
        panel.grid.minor = element_line(linewidth = .45, linetype = "dashed"),
        
        plot.background = element_rect(fill = "#d9e3f1", color = NA),
        plot.margin = margin(20, 20, 20, 20)
    )
```

### Boxplot 2

plot for boxplot 2

```{r}
# plot 2 ---------------------

df |>
    ggplot(aes(Group, Shannon)) +
    
    geom_point(
        aes(fill = Patient),
        position = position_jitterdodge(jitter.width = .15, dodge.width = .5),
        shape = 21, size = 2, stroke = .15, color = "white"
    ) +
    
    geom_boxplot(aes(fill = Patient), alpha = .5, width = .5) +
    
    theme_minimal() +
    
    theme(
        plot.margin = margin(20, 20, 20, 20)
    )
```
