---
title: "Heatmap"
format:
  html:
   fig-width: 13
   fig-height: 15
execute: 
  warning: false

---

A heatmap is a graphical representation of data where the individual values contained in a matrix are represented as colors.


## Table 1 (Data format)

This is an example of a data format containing the necessary information for heatmap visualizations. The table includes mock data specifically generated for this purpose.


```{r}

library(data.table)
library(reactable)

ann <- fread("../inst/data/ann.txt") 


reactable(
  
  ann, 
  defaultPageSize = 5,
  theme = reactableTheme(
    backgroundColor  = "#d9e3f1"
  )
)
```


## Table 2 (Data format)

This is an example of a data format containing the necessary information for heatmap visualizations. The table includes mock data specifically generated for this purpose.


```{r}

library(data.table)
library(reactable)

df <- fread("../inst/data/gene_counts.txt") 

mm = df[, ann$Sample, with = FALSE] |> setDF(rownames = df$GeneID) |> as.matrix()

zmat = mm |> t() |> scale(scale = TRUE, center = TRUE) |> t()

reactable(
  
  zmat, 
  defaultPageSize = 5,
  theme = reactableTheme(
    backgroundColor  = "#d9e3f1"
  )
)
```

## Heatmap plot

The heatmap() function is a built-in feature of R, providing a powerful tool for generating high-quality heatmaps from matrices. It includes statistical functionalities to normalize input data, perform clustering algorithms and visualize the results


```{r}

# libraries -------
library(ComplexHeatmap)
library(ggplotify)
library(ggplot2)

ht = Heatmap(
    zmat, name = "Z-score",
    
    border = TRUE,
    
    clustering_distance_columns = "euclidean",
    clustering_distance_rows = "euclidean",
    
    clustering_method_columns = "ward.D2",
    clustering_method_rows = "ward.D2",
    
    column_split = ann$Group1,
    row_split = 2,
    
    row_names_gp = gpar(fontsize = 4)
)


ht |> 
    draw(merge_legends = TRUE, background = "transparent") |>
    grid.grabExpr() |>
    as.ggplot()

```
